---
title: "Lecture1"
author: "Elaine M. Wright"
format: html
editor: visual
---

# 09.07.22 - Lecture #1

New class partners: Jake and Murry

## Functions

-   Functions are denoted with ( )

```{r}
# Read this as: call the sin function and pass it the argument 1
sin(1)

# This works but it is considered bad style
sin (1)
```

-   Infix operators - "secret functions"

    -   do not require parenthesis

    -   (+), (\^), (\*)

-   Logical operators (infix) - compares things and returns as T or F

```{r}
1 == 1
1 == 2
1 < 3
1 <= 4
1 != 1
!TRUE
```

All of this R code is currently not in memory, this is because none of the code is currently assigned to an object

## Assignment

-   R will autoprint for you by just running the variable in the console

-   R allows '=' to be used for assignment like in python, but we will not use this for assignment due to another function of the '=' within R

```{r}
x <- 1/40
x # this is an example of autoprinting 
x + 1

y = 1/40

# Assignment can also work in this order, not recommended
1 -> d 
```

## Naming Objects

-   R is **case sensitive**

-   avoid capital letters for purpose of this class

```{r}
tree.heights # snake case
treeHeights # camel case
tree_heights #?
tree.heights.v2 <- 4
.tree.heights <- 3 #hidden variable

print(.tree.heights)
```

## Legal & Illegal Naming of Objects

```{r}
min_height <- 2
max.height <- 9
_age <- 5       # Illegal variable name
.mass <- 6      # Hidden variable
MaxLength <- 7
min-length <- 1 # This becomes a function
2widths <- 12   # Illegal variable name
celcius2kelvin <- 10
```

## Vectors

-   Colons ( : ) denotes a sequence of integers from x to x

```{r}
1:5
vec1 <- 1:5
vec2 <- 6:10
vec1 + vec2
vec1 * vec2
```

-   We use vector functions whenever possible

-   R also compiles per line instead of all at once like C does

## Managing the Envrionment

```{r}
ls() # list objects in environment
rm(celcius2kelvin)
rm(list = ls()) # argument named list is given the value ls()
```
