---
title: "Hmk_05"
author: "Elaine M. Wright"
format: gfm
editor: visual
---

## Installing Packages:

```{r, message=FALSE}
library(nycflights13)
library(tidyverse)

#Test to make sure package loaded successfully
nycflights13::flights
```

## Filtering:

A plot of air time as a function of distance for all flights that meet the following criteria:

-   originate from LaGuardia airport ("LGA")
-   departed on the 16th of the month
-   have a flight distance of less than 2000

```{r}
filtered <- filter(flights, origin == "LGA", day == 16, air_time < 2000)
ggplot(filtered, aes(x=distance, y=air_time)) +
  geom_point()
```

## Dealing With NA's:

A new data frame of all of the rows of `flights` that have values for *both* `arr_time` and `dep_time`, where neither value is `NA`.

```{r}
filter(flights, arr_time != "NA" & dep_time != "NA")
```

```{r}
# UPDATED BASED ON FEEDBACK:
# New way to filter NA values:

no_na <- filter(flights, !(is.na(air_time)|is.na(distance)))
no_na
```

## Filtering NA's:

Preventing NA values from appearing in the first place can be done with `na.rm = TRUE`

## Adding Columns:

A new data frame was selected to show the airline carrier, the air time, and the distance of the flight. A new column was added to the data frame using `mutate()` showing the average flight speed based `air_time` and `distance`. A histogram plot was made and shows the average flight speed for every air line carrier.

```{r}
average <- select(flights, carrier, air_time, distance)
mutate(average,
       speed <- distance / air_time * 60)
# Previously the air time was not multiplied by 60, multiplying by 60 actually gives the speed in mph
```

```{r}
ggplot(average, aes(x=speed)) +
  geom_density()+
  xlab("Av. Speed (mph)")
```
